---
title: "Blocking Sockets in Two R Sessions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Blocking Sockets in Two R Sessions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Blocking calls wait until the peer performs the matching operation. This
example uses two distinct R sessions on the same machine:

- **Session 1** starts a server and blocks in `socket_accept()` and then
  `socket_receive()`.
- **Session 2** connects to the server and sends the message that releases
  Session 1.

Choose a free TCP port, such as `45678`, and run Session 1 first.

## Session 1: blocking server

Run this code in the first R session:

```{r blocking-server, eval=FALSE}
library(socketR)

server <- socket_create("inet", "stream")
client <- NULL
on.exit({
  if (!is.null(client) && socket_is_open(client)) socket_close(client)
  if (socket_is_open(server)) socket_close(server)
}, add = TRUE)

socket_reuse_address(server, TRUE)
socket_bind(server, "127.0.0.1", 45678L)
socket_listen(server)
cat("Waiting for Session 2...\n")

# This call blocks until Session 2 connects.
client <- socket_accept(server)
cat("Session 2 connected; waiting for data...\n")

# This call blocks until Session 2 sends bytes.
# connection loop starts here; this call blocks until Session 2 sends the bytes.
payload <- socket_receive(client, n = 1024L)
cat("Received:", rawToChar(payload), "\n")
socket_send(client, "ack from Session 1")

socket_close(client)
```

## Session 2: blocking client

After Session 1 prints that it is waiting, run this code in a second R
session:

```{r blocking-client, eval=FALSE}
library(socketR)

client <- socket_create("inet", "stream")
on.exit(if (socket_is_open(client)) socket_close(client), add = TRUE)

# This call blocks until Session 1 is listening and accepts the connection.
socket_connect(client, "127.0.0.1", 45678L)

# communication loop starts here; this call blocks until Session 1 receives the bytes.
socket_send(client, "hello from Session 2")

# This call blocks until Session 1 sends the acknowledgement.
ack <- socket_receive(client, n = 1024L)
cat("Received:", rawToChar(ack), "\n")
```

Both sockets are blocking by default. To switch either session to
nonblocking behavior, call `socket_set_blocking(client, FALSE)` and use
`socket_poll()` before receiving data. Always close handles explicitly in
long-running sessions; the external-pointer finalizer is a last-resort
cleanup mechanism.
